The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}

### 2017-06-12 GRB111117A.tex
Address Andrejas comments and drop letter format
Closes #17

There is mounting evidence that short-duration $\gamma$-ray bursts
come from the merger of NSs, either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

@@

%So far, the electromagnetic signal from sGRBs are the only means to accurately
% localize NS mergers and holds the promise for detection of an associated
% gravitational wave signal \citep{Ghirlanda2016}. From the burst and following
% afterglow, modeling the properties and mapping the environments additionally
% allows insights into the phenomenon itself.

@@

In this \emph{Letter} we present a spectrum of the host galaxy of the short
GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This
value is significantly higher than the previously estimated redshift based on
photometric studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest
frame properties based on this new distance compared to previous analyses and
revisit the host properties derived from the new solution to the spectral energy distribution (SED) fit.
%While no optical afterglow was detected, the excellent localization from a
%detection of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also
%allows us to discuss the positioning and environmental properties of this
%remarkably distant short GRB.
Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology parameters provided by
\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
and $\Omega_m = 0.307$. All magnitudes are given in the AB system.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs4.pdf}
	\caption{
	Imaging of the field of GRB~111117A with the X-shooter slit overlaid. The slit
	position represents 4 epochs of spectroscopic observations taken at similar
	position angles. The image is the FORS2 $R$-band image for which the photometry
	is shown in Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as
	derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. The
	insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The location
	of a bright skyline is marked by the blue box. The location of the emission
	line is indicated with the red ellipse. Because the host is observed in
	nodding-mode, negative images of the emission line appear on both sides in the
	spatial direction.
	}
	\label{fig:spec_setup}
\end{figure}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). X-shooter covers the wavelength range from 3000~\AA~to
24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
spectroscopic arms. We carried out the bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration using the
VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}.
%Because the spectra are curved across each detector, a rectification algorithm
%is employed which introduces correlations between neighboring pixels. We select
%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
%degree of correlation while conserving the maximal resolution.
The observations are combined and extracted using scripts described in Selsing
et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. 
%The signal-to-noise ratio of the continuum in the near-infrared arm is too low
%to use the optimal extraction algorithm \citep{Horne1986} and therefore the
%extraction is carried out with a simple aperture.
An overview of the spectroscopic observations is given in
Table~\ref{tab:spec_overview}, and the slit position is shown in
Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in Fig.~\ref{fig:SED}.

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines belonging to \lya, \oii, \hb, \oiii$\lambda$5007, and \ha. 
\hb~is detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
in the insert in Fig.~\ref{fig:spec_setup}.  \ha~is only visible in the first
epoch due to the $K$-band blocking filter used for the remainder observations.
The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
significant velocity structure along the slit. A drop in the continuum to the
blue of the \lya~line further supports the inferred redshift.

%The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
% profiles shows an asymmetric "double-horned" profile, indicating that we are
% seeing a galaxy with a large degree of coherent rotational motion relative to
% the line-of-sight. If we assume that we are viewing a spiral galaxy exactly
% edge-on, this is a measure of the rotational velocity of the gas. If we assume
% that the spectral resolution and the turbulent width of the lines are
% negligible compared to the rotational velocity, we can based on the projected
% size of the source and the width of the lines, put a constraint on the
% dynamical mass of the galaxy. Based on the physical scale from extent and the
% velocity with, we infer M$_\text{dyn} \gtrsim 10^{10.8}$ M$_\odot$. Because we
% are viewing the host at an angle, this value is an upper limit, which should
% be divided by $\sin(i)$, where $i$ is the inclination relative to edge-on.

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a 
galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at 
21\,000~\AA~where the night sky is very bright. In addition, several bright sky-lines 
are superposed on the line, making an accurate estimate of the \ha-flux difficult. 
We obtain a limit on the SFR by integrating the part of \ha~free of contamination and 
correcting for the missing flux using the line shape. After converting the 
\citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass function using
\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$. 
From the SED-fit (Sect.~\ref{SED}), and the detection of \lya, the host is
constrained to contain very little or no dust, although the presence of Lya does
not exclude dust. Therefore we do not apply a dust-correction to the measured
\ha~flux here.

%Secondly, based on the width of \oiii, an aperture covering the line
% is integrated over, where synthetic sky lines \citep{Noll2012, Jones2013} have
% been masked and interpolated over. From the integrated \ha-line, we estimate
% SFR = $18 \pm 3$ M$_\odot$ yr$^{-1}$.

%\begin{figure}
%	\centering
%	\includegraphics[width=9cm]{figures/OIII_img.pdf}
%	\caption{2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
%	\label{fig:line}
%\end{figure}

%A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
% "darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}. Using the optical
% afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
% interpolated and evaluated at the time of the non-detection. We find
% $\beta_{OX} < 0.79$, consistent with what was reported in \citet{Sakamoto2013}.

@@

All data were reduced, analyzed and fitted in a similar manner as described in
detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
reduction which includes bias/flat-field correction, de-fringing (if necessary),
sky-subtraction, and stacking of individual images. The photometry of the host
was calibrated relative to field stars from the SDSS and 2MASS catalogs in the
case of $grizJHK_{\mathrm{s}}$ filters.
%For the $R$ and $I$-band photometry, we used the color transformations of
%Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
We convert all magnitudes into the AB system, and correct for a Galactic 
foreground of $E_{B-V}=0.027~\mathrm{mag}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Redshift, line-fluxes, SFR, and Mass}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
\citep{Evans2009}. 
%The X-ray afterglow light curve can be fit with a single power-law decay with an
%index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
%analysis using data obtained in photon counting (PC) mode in the widest time epoch where
%the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
%(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
%data) to prevent spectral changes that can affect the X-ray column density
%determination. 
The obtained spectrum is well described by an absorbed power-law
model and the best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
an intrinsic equivalent hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a solar abundance and a Galactic $N_{\mathrm{H}}$ in
the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is
securely classified as belonging to the short class of GRBs. Because the
observed classification indicators, $T_{90}$ and hardness ratio, do not depend
strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not
change this designation.
%The intrinsic luminosity is higher as demonstrated in the X-ray light curve
%(Fig. \ref{fig:sxray_lightcurve}), but it is still is sub-luminous compared to
%the majority of long GRBs.
\citet{Bromberg2013} investigated the degree to which the long and short
population distributions overlap and quantified the certainty in the class
membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
percent probability of being a sGRB. Compared to the other two sGRB candidates
at high redshift, GRB~060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7
\lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB~090426
\citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
per cent), the certainty in class membership for GRB~111117A is much higher.
This means that GRB~111117A happened when the universe was younger by 1.8 Gyr
compared to any other non-collapsar GRB ever detected. This number is 3.2 Gyr
for the next-highest spectroscopic redshift. This has implications for the
$r$-process element enrichment timescale of the universe.

%Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
%intermediate duration bursts. This comes from both events having very soft
%spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
%bursts are very clearly related in their properties to lGRBs
%\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
%mergers.

%\begin{figure}
%	\centering
%	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
%	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs. The grey shaded region is a compilation of long GRB lightcurves from \todo{reference} where the color represents density and the light blue lines are other short GRB lightcurves from \todo{reference} for which redshifts has been determined. Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
%	\label{fig:sxray_lightcurve}
%\end{figure}

@@

We show the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare
with the distributions of complete samples of both long and short GRBs. The lGRB
sample is from \citet{Arcodia2016} and the sGRB sample is from
\citet{DAvanzo2014a}. 17 out of the 99 long bursts do not have measured redshifts, and
likewise for 5 out of 16 in the short sample. Bursts without measured redshifts have been
excluded for both groups. GRB~111117A occupies a unique position in
Fig.~\ref{fig:NH_z} with the highest $N_H$ of all short bursts. Additionally the
upper limit at $z = 2.609$ belongs to GRB~090426 and does not belong in the
short sample. The short sample, excluding GRB~111117A, is located at low
redshifts ($z < 1$) and is found to populate a similar column density
environment to lGRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
hydrogen column density for GRB~111117A seems to follow the trend with increasing $N_H$
as a function of redshift as found for the lGRB afterglows \citep{Arcodia2016}.

%For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
%\citep{Lyman2017}, which is a challenge to reconcile with the relatively large
%offset from the host center derived in \citet{Margutti2012} and
%\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
%burst position that we don’t see. Along with the absence of dust, the large
%offset from the host center indicates that the high $N_H$ arises because the
%density in the GRB surrounding is high, or because the light from the afterglow
%traverses a region of star formation.

@@

%From the clear host association, GRB~111117A does not belong to the hostless
%class of GRBs \citep{Berger2010a} and because the host exhibits emission lines
%this is indicative of a population of relatively young stars.
As the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is a
late-type galaxy and is entirely consistent in terms of stellar mass and stellar
age with the general population of sGRB hosts ($\left\langle M _*
\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
and sSFR are entirely within the range expected for the hosts of sGRBs
\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
typical SFR for sGRB host galaxies \citep{Berger2014} and more similar to
the SFR found in the hosts of lGRBs at a corresponding redshift
\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
star forming galaxy would exhibit weaker emission lines, thus making the
redshift harder to determine. Additionally, it is natural to expect some
evolution in the hosts of sGRBs with redshift as illustrated in Sect.~\ref{restnH}.

%The detection of \lya~is consistent with the SED-inferred absence of dust,
%despite the moderate stellar age which would suggest the opposite. The centroid
%of the Ly$\alpha$ emission is found to be redshifted by $\sim$~220~km~s$^{-1}$
%with respect to systemic, which is similar to what is found for long GRB hosts
%\citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.

@@

A fraction of the bursts missing redshift are host-less and is therefore likely
at moderate redshifts \citep{Tunnicliffe2014}, but should some of the remainder
be at high redshift, the missed fraction will increase. If we assume that
\textit{all} the bursts that are missing redshifts are at high-$z$ and missed
due to host faintness, 9 out of 14 hosts in the complete sample (64 per cent)
would be missed at $z = 2.211$. This serves as an upper limit on the fraction of
missed burst redshifts at high-$z$. The corresponding fraction of missed bursts
is 0.55 (22 out of 39 hosts) using the full sample from \citet{Berger2014}.

@@

In practice most NS-NS binaries will be eccentric at formation because of the SN
natal kicks, although eccentric systems normally circularise first. For more
eccentric orbits, the coalescence timescale decreases, leading to a decrease in
the constraint on the initial separation and larger initial separations. As
noted by \citet{Postnov2014}, it takes eccentricities > 0.6 to significantly
shorten the merger time. 
%It could also be a BH-NS system, and as the merger time scales roughly as $M^3$
%the mass here almost as big an effect as the separation ($a^4$).

@@

In this \emph{Letter}, we have provided a revised, spectroscopic redshift
measurement for the short GRB~111117A based on host galaxy emission lines
setting it at $z = 2.211$. This value supersedes the previous photometric
redshift of \citet{Margutti2012} and \citet{Sakamoto2013}. 
%The erroneous redshift estimate of previous authors is attributed to a
%discrepancy in the measured $z$-band magnitude.

@@

Although a single burst carries little leverage in terms of constraining the
redshift distribution of sGRB, the high redshift of GRB~111117A needs to be
accommodated in progenitor models. A lognormal delay time model predicts a very
low volumetric density of bursts at $z = 2.211$, whereas a power law delay time
model peaks near GRB~111117A. If more sGRBs are at similarly high redshifts, but
are missed due to the faintness of their hosts, a lognormal delay time model
will be disfavored. Compared to a sample of sGRB hosts, GRB~111117A is more
luminous than 80 per cent of the sample with measured luminosities.
Conservatively, for 64 per cent of sGRB hosts, we would be unable to determine a
redshift should they be at a similar redshift as GRB~111117A.

@@

\newpage
%\appendix

>>>>

There is mounting evidence that short-duration $\gamma$-ray bursts
come from the merger of NSs, either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

@@

So far, the electromagnetic signals from sGRBs are the only means to accurately
localize NS mergers and hold the promise for detection of an associated
gravitational wave signal \citep{Ghirlanda2016}. From the burst and following
afterglow, modeling the properties and mapping the environments additionally
allows insights into the phenomenon itself.

@@

We here present a spectrum of the host galaxy of the short
GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This
value is significantly higher than the previously estimated redshift based on
photometric studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest
frame properties based on this new distance compared to previous analyses and
revisit the host properties derived from the new solution to the spectral energy distribution (SED) fit.
%While no optical afterglow was detected, the excellent localization from a
%detection of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also
%allows us to discuss the positioning and environmental properties of this
%remarkably distant short GRB.
We use the $\Lambda$CDM cosmology parameters provided by
\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
and $\Omega_m = 0.307$. All magnitudes are given in the AB system.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs.pdf}
	\caption{
	Imaging of the field of GRB~111117A with the X-shooter slit overlaid. The slit
	position represents 4 epochs of spectroscopic observations taken at similar
	position angles. The image is the FORS2 $R$-band image for which the photometry
	is shown in Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as
	derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. 
%	The insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The
%	location of a bright skyline is marked by the blue box. The location of the
%	emission line is indicated with the red ellipse. Because the host is observed
%	in nodding-mode, negative images of the emission line appear on both sides in
%	the spatial direction.
	}
	\label{fig:spec_setup}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/OIII_img.pdf}
	\caption{2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
	\label{fig:line}
\end{figure}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). X-shooter covers the wavelength range from 3000~\AA~to
24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
spectroscopic arms. We carried out the bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration using the
VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}.
Because the spectra are curved across each detector, a rectification algorithm
is employed which introduces correlations between neighboring pixels. We select
a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
degree of correlation while conserving the maximal resolution.
The observations are combined and extracted using scripts described in Selsing
et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. 
The signal-to-noise ratio of the continuum in the near-infrared arm is too low
to use the optimal extraction algorithm \citep{Horne1986} and therefore the
extraction is carried out with a simple aperture covering the entire host.
An overview of the spectroscopic observations is given in
Table~\ref{tab:spec_overview}, and the slit position is shown in
Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in Fig.~\ref{fig:SED}.

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines belonging to \lya, \oii, \hb, \oiii$\lambda$5007, and \ha. 
\hb~is detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
in the insert in Fig.~\ref{fig:line}.  \ha~is only visible in the first
epoch due to the $K$-band blocking filter used for the remaining observations.
The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
significant velocity structure along the slit. A drop in the continuum to the
blue of the \lya~line further supports the inferred redshift.

The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
 profiles shows an asymmetric "double-horned" profile, indicating that we are
 seeing a galaxy with a large degree of coherent rotational motion relative to
 the line-of-sight. If we assume that we are viewing a spiral galaxy exactly
 edge-on, this is a measure of the rotational velocity of the gas. If we assume
 that the spectral resolution and the turbulent width of the lines are
 negligible compared to the rotational velocity, we can based on the projected
 size of the source and the width of the lines, put a constraint on the
 dynamical mass of the galaxy. Based on the physical scale from extent and the
 velocity with, we infer M$_\text{dyn} \gtrsim 10^{10.8}$ M$_\odot$. Because we
 are viewing the host at an angle, this value is an upper limit, which should
 be divided by $\sin(i)$, where $i$ is the inclination relative to edge-on.

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a
galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed
at 21\,000~\AA~where the night sky is very bright. In addition, several bright
sky-lines are superposed on the line, making an accurate estimate of the
\ha-flux difficult. We obtain a limit on the SFR by integrating the part of
\ha~free of contamination and correcting for the missing flux using the line
shape. After converting the \citet{Kennicutt1998} relation to a
\citet{Chabrier2003} initial mass function using \citet{Madau2014}, we derive a
limit of $SFR > 7 M_\odot$~yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and
the detection of \lya, the host is constrained to contain very little or no
dust, although the presence of Lya does not exclude dust. Therefore we do not
apply a dust-correction to the measured \ha~flux here. Secondly, based on the
width of \oiii, an aperture covering the line is integrated over, where
synthetic sky lines \citep{Noll2012, Jones2013} have been masked and
interpolated over. From the integrated \ha-line, we estimate SFR = $18 \pm 3$
M$_\odot$ yr$^{-1}$.

A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
"darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}. Using the optical
afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.

@@

All data were reduced, analyzed and fitted in a similar manner as described in
detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
reduction which includes bias/flat-field correction, de-fringing (if necessary),
sky-subtraction, and stacking of individual images. The photometry of the host
was calibrated relative to field stars from the SDSS and 2MASS catalogs in the
case of $grizJHK_{\mathrm{s}}$ filters.
For the $R$ and $I$-band photometry, we used the color transformations of
Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
We convert all magnitudes into the AB system, and correct for a Galactic 
foreground of $E_{B-V}=0.027~\mathrm{mag}$.

@@

@@

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
\citep{Evans2009}. 
The X-ray afterglow light curve can be fit with a single power-law decay with an
index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
analysis using data obtained in photon counting (PC) mode in the widest time epoch where
the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
data) to prevent spectral changes that can affect the X-ray column density
determination. 
The obtained spectrum is well described by an absorbed power-law
model and the best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
an intrinsic equivalent hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a solar abundance and a Galactic $N_{\mathrm{H}}$ in
the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is
securely classified as belonging to the short class of GRBs. Because the
observed classification indicators, $T_{90}$ and hardness ratio, do not depend
strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not
change this designation.
The intrinsic luminosity is higher as demonstrated in the X-ray light curve
(Fig. \ref{fig:sxray_lightcurve}), but it is still is sub-luminous compared to
the majority of long GRBs.
\citet{Bromberg2013} investigated the degree to which the long and short
population distributions overlap and quantified the certainty in the class
membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
percent probability of being a sGRB. Compared to the other two sGRB candidates
at high redshift, GRB~060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7
\lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB~090426
\citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
per cent), the certainty in class membership for GRB~111117A is much higher.
This means that GRB~111117A happened when the universe was younger by 1.8 Gyr
compared to any other non-collapsar GRB ever detected. This number is 3.2 Gyr
for the next-highest spectroscopic redshift. This has implications for the
$r$-process element enrichment timescale of the universe.

Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
intermediate duration bursts. This comes from both events having very soft
spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
bursts are very clearly related in their properties to lGRBs
\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
mergers.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs. The grey shaded region is a compilation of long GRB lightcurves from \todo{reference} where the color represents density and the light blue lines are other short GRB lightcurves from \todo{reference} for which redshifts has been determined. The thin green line is the light-curve using $z = 1.3$ and the think green line is for $z = 2.211$. Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
	\label{fig:sxray_lightcurve}
\end{figure}

@@

We show the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare
with the distributions of complete samples of both long and short GRBs. The lGRB
sample is from \citet{Arcodia2016} and the sGRB sample is from
\citet{DAvanzo2014a}. 17 out of the 99 long bursts do not have measured redshifts, and
likewise for 5 out of 16 in the short sample. Bursts without measured redshifts have been
excluded for both groups. GRB~111117A occupies a unique position in
Fig.~\ref{fig:NH_z} with the highest $N_H$ of all short bursts. Additionally the
upper limit at $z = 2.609$ belongs to GRB~090426 and does not belong in the
short sample. The short sample, excluding GRB~111117A, is located at low
redshifts ($z < 1$) and is found to populate a similar column density
environment to lGRBs at similar redshifts \citep{DAvanzo2014a}. 
The inferred hydrogen column density for GRB~111117A seems to follow the trend
with increasing $N_H$ as a function of redshift as found for the lGRB afterglows
\citep{Arcodia2016}. This is similar to what is found by \citet{Kopac2012,
	Margutti2013}, that $N_H$ seems to be comparable for long and short GRBs when
compared at similar redshifts.

For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
\citep{Lyman2017}, which is a challenge to reconcile with the relatively large
offset from the host center derived in \citet{Margutti2012} and
\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
burst position that we don’t see. Along with the absence of dust, the large
offset from the host center indicates that the high $N_H$ arises because the
density in the GRB surrounding is high, or because the light from the afterglow
traverses a region of star formation.

@@

From the clear host association, GRB~111117A does not belong to the hostless
class of GRBs \citep{Berger2010a} and because the host exhibits emission lines
this is indicative of a population of relatively young stars.
As the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is a
late-type galaxy and is entirely consistent in terms of stellar mass and stellar
age with the general population of sGRB hosts ($\left\langle M _*
\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
and sSFR are entirely within the range expected for the hosts of sGRBs
\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
typical SFR for sGRB host galaxies \citep{Berger2014} and more similar to
the SFR found in the hosts of lGRBs at a corresponding redshift
\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
star forming galaxy would exhibit weaker emission lines, thus making the
redshift harder to determine. Additionally, it is natural to expect some
evolution in the hosts of sGRBs with redshift as illustrated in Sect.~\ref{restnH}.

The detection of \lya~is consistent with the SED-inferred absence of dust,
despite the moderate stellar age which would suggest the opposite. The centroid
of the Ly$\alpha$ emission is found to be redshifted by $\sim$~220~km~s$^{-1}$
with respect to systemic, which is similar to what is found for long GRB hosts
\citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.

@@

A fraction of the bursts missing redshift are host-less and is therefore likely
at moderate redshifts \citep{Tunnicliffe2014}, but should some of the remainder
be at high redshift, the missed fraction will increase. If we assume that
\textit{all} the bursts that are missing redshifts are at high-$z$ and missed
due to host faintness, 9 out of 14 hosts in the complete sample (64 per cent)
would be missed at $z = 2.211$. This serves as an upper limit on the fraction of
missed burst redshifts at high-$z$. The corresponding fraction of missed bursts
is 0.55 (22 out of 39 hosts) using the full sample from \citet{Berger2014}. This
indicates that we could be missing a significant high-$z$ sGRBs.

@@

In practice most NS-NS binaries will be eccentric at formation because of the SN
natal kicks, although eccentric systems normally circularise first. For more
eccentric orbits, the coalescence timescale decreases, leading to a decrease in
the constraint on the initial separation and larger initial separation constraints. As
noted by \citet{Postnov2014}, it takes eccentricities > 0.6 to significantly
shorten the merger time. 
%It could also be a BH-NS system, and as the merger time scales roughly as $M^3$
%the mass here almost as big an effect as the separation ($a^4$).

@@

We have here provided a revised, spectroscopic redshift measurement for the
short GRB~111117A based on host galaxy emission lines setting it at $z = 2.211$.
This value supersedes the previous photometric redshift of \citet{Margutti2012}
and \citet{Sakamoto2013}. %The erroneous redshift estimate of previous authors is
attributed to a discrepancy in the measured $z$-band magnitude.

@@

Although a single burst carries little leverage in terms of constraining the
redshift distribution of sGRB, the high redshift of GRB~111117A needs to be
accommodated in progenitor models. A lognormal delay time model predicts a very
low volumetric density of bursts at $z = 2.211$, whereas a power law delay time
model peaks near GRB~111117A. If more sGRBs are at similarly high redshifts, but
are missed due to the faintness of their hosts, a lognormal delay time model
will be disfavored. Compared to a sample of sGRB hosts, GRB~111117A is more
luminous than 80 per cent of a complete sample of sGRB hosts with measured
luminosities. Conservatively, for 64 per cent of sGRB hosts, we would be unable
to determine a redshift should they be at a similar redshift as GRB~111117A.
This could indicate that a significant fraction of sGRBs are at high $z$.

@@
